### 2018-05-22 chapters/02-background/dlmf.tex

\section{Digital Library of Mathematical Functions}\label{sec:dlmf}
\glsreset{dlmf}\glsreset{drmf}\glsreset{opsf}
The \gls{nist} \gls{dlmf}~\cite{NIST:DLMF:Paper} is the digital version of the \textit{NIST Handbook of Mathematical Function}. Internally, the \gls{dlmf} websites are \LaTeX{} files and all equations are numbered as usual. Therefore, we reference to a formula in the \gls{dlmf} with the given equation number.

The \gls{nist} \gls{drmf}~\cite{DRMF:14} is an outgrowth of the \gls{dlmf} and wants to create a digital compendium of mathematical formulae for \gls{opsf}. One of the main purposes is to facilitate interaction among a community of mathematicians and scientists~\cite{DRMF:15}. The \gls{drmf} also aims at expandability and interactivity. That was also a reason for this translation project, because a goal was to provide a one-click translation on the formulae in the \gls{drmf}.

Together, the \gls{dlmf} and \gls{drmf} build a comprehensive compendium for \gls{opsf} and they are therefore our main references in this thesis. A definition of a function in any \gls{cas} is therefore called different, if it differs from the corresponding \gls{dlmf}/\gls{drmf} definition.