### 2018-11-27 ReviewerResponse.txt
Improve response to reviewers (close #19)

In the following we respond specifically to the reviewers comments.

Reviewer 1:

As suggested, we made several changes to improve the structure of the paper.  
In summary, we have :
(1) added subsection headings "Different Sets of Defined Functions", 
"Positions of Branch Cuts", "Insufficient Semantic Information", and 
"Ambiguous Expressions" in Section 3;
(2) gave an overview of the concept of the translator in Section 4;
(3) removed unnecessary Algorithm 1 in Section 5 as suggested, because 
it was just a simplified pre-version of Algorithm 2;
(4) added the subsection heading "Problems with the Look Ahead Approach" in 
Section 5 for a better outline; and 
(6) further explained the different states of Algorithm 3 (now Algorithm 2) 
by refer to the lines in the descriptions. We also explained how a power 
after a macro is handled in more detail and why optional parameters needs 
to be handled differently to other arguments.

We also gave further detailed descriptions about the PoM-Tagger, 
namely its parsing process and lexicon files.

We also added LaTeXML to the related work section and gave a brief 
explanation with cited research (Schubotz et al., 2018, now included 
in our references), in order to explain why we did not choose MathML 
as an intermediate data format for performing translations. In that 
citation, we described detailed testing that we performed using LaTeXML 
and other state of the art LaTeX to MathML conversion tools. In that
reference, we discovered several critical problems which could occur 
when translations would be performed using MathML as an intermediate
data format.

Note that we did not add further comparisons with LaTeXML to our 
evaluations section, because current CAS are unable to understand MathML 
which contains special functions. Thus we can only compare translations 
without function calls or implement a solution which overcomes this 
problem, in order to make a decent comparison. Because of the great amount 
of problems which would need to be overcome, a full solution translation 
implementation using MathML is out of the scope of our current project. 
This is now all explained in detail in the new version of the manuscript.

Reviewer 2:

For clarification, throughout the manuscript, we decided to change the 
wording "Word Processor" to "Document Preparation System." Since this wording
occurred as part of the title, we did not add that we focus on LaTeX to the 
title and hope that the Abstract points this out in an acceptable manner.

Also, instead of creating an Appendix, we decided to add short paragraphs 
which explain the purpose of the Sections 4 and 6. We hope 
this helps the reader to decide whether he can skip the section, e.g., in 
case of his lack of interest in Maple "hackings", or not.  In the new 
version of the paper, we implemented all suggested minor changes as well as 
adding the idea of improved numerical tests by using the mentioned cut 
region detection to the future work section.