### 2018-11-29 ReviewerResponse.txt
updated ReviewerResponse.txt

In the following we respond specifically to the reviewers comments.

Reviewer 1:

As suggested, we made several changes to improve the structure of the paper.  
In summary, we have :
(1) added the subsection headings "Different Sets of Defined Functions", 
"Positions of Branch Cuts", "Insufficient Semantic Information", and 
"Potentially Ambiguous Expressions" in Section 3; 
(2) gave an overview of the concept of the translator in Section 4;
(3) as suggested, we removed the unnecessary Algorithm 1 in Section 5 
because it was just a simplified pre-version of Algorithm 2;
(4) added the subsection heading "Problems with the Look Ahead Approach" in 
Section 5 for a better outline; and 
(5) further explained the different states of Algorithm 3 (now Algorithm 2) 
by referring to specific lines in the algorithm.

We explained how a power after a macro is handled in more detail and why 
optional parameters need to be handled differently than other arguments.

We gave further detailed descriptions concerning the PoM-Tagger, namely its 
parsing process and lexicon files.

We added LaTeXML to the related work section and gave a brief 
explanation with cited research (Schubotz et al., 2018, now included 
in our references), in order to explain why we did not choose MathML 
as an intermediate data format for performing translations. In that 
citation, we described detailed testing that we performed using LaTeXML 
and other state-of-the-art LaTeX to MathML conversion tools. In that
reference, we discovered several critical problems which could occur 
when translations would be performed using MathML as an intermediate data format.

Note that we did not add further comparisons with LaTeXML to our 
evaluations section, because current CAS are unable to understand MathML 
which contains special functions. Thus, in order to make a decent 
comparison, we can only compare translations without function calls or 
implement a solution which is able overcome these problems.  Because of 
the problems which would need to be overcome, a full solution translation 
implementation using MathML is out of the scope of our current project. 
This is now all explained in detail in the new version of the manuscript.

Reviewer 2:

For clarification, throughout the manuscript, we decided to change the 
wording "Word Processor" to "Document Preparation System," and 
"Representations of Special Functions" to "Expressions involving Special 
Functions." Since this wording occurred as part of the title, we did not 
add that we focus on LaTeX to the title and hope that the Abstract points 
this out in an acceptable manner.

Instead of creating an Appendix, we decided to add short paragraphs 
which explain the purpose of Sections 4 and 6. We hope this helps the reader 
to decide whether he can skip the section, e.g., in case of his lack of 
interest in Maple "hackings", or not.  In the new version of the paper, 
we implemented all suggested minor changes as well as adding the idea of 
improved numerical tests by using the mentioned cut region detection to 
the future work section.

To address the weakness in precision, we added an enhanced definition for 
appropriate and inappropriate translations. Furthermore, we have rewritten
the paragraph about alternative translations for the arccotangent function 
on Page 6 (now Page 7).